Hadith Narrator Chain Extraction using Arabic Morphological Analysis

### 2011-02-25 tex/hadith.short.tex
last change

\begin{document}
\maketitle
\begin{abstract}
A hadith in Islamic literature is a narration from the 
prophet Mohammad related by multiple narrators.
Establishing the authenticity of a hadith is an important task
in Islamic studies. 
Based on them Islamic scholars issue rules that affect the life
of people around the world. 
The literature lacks an automated exhaustive mechanism of 
authentication checks. 
This is currently manual work prone to error and almost
impossible to complete in the life time of a scholar due to the
large literature available. 
In this paper, we present a tool that uses 
an Arabic application-specific morphological analyzer 
to successfully automate the
analysis of three books of hadith selected 
arbitrarily and abstract 
each book into a graph of narrators related to each other. 
\end{quote}
\end{abstract}

\section{Introduction}

\transfalse
\begin{figure}[tb]
\center{
\resizebox{.9\columnwidth}{!}
{ \input{figs/exhadith.pdftex_t}}
\caption{Hadith abstraction example.}
\label{f:exhadith}
}
\end{figure}
% the FSMs for the three words
\transtrue

A \RL{.hady_t}~\footnote{In this document, 
we use the default ArabTeX transliteration style ZDMG.}
is a narration related to the prophet Mohammad
through a \RL{sanad} or a sequence of narrators. 
The collections of traditions are the second source of
jurisprudence after the \RL{qor'An} for all Islamic schools of thought. 
Figure~\ref{f:exhadith} shows an example \noArRL{.hady_t} in 
Arabic with its transliteration and translation. 
We show proper names in boxes connected
to form complex names of narrators. 
For example, 
\noTrRL{qtybT} is the first name 
of narrator $n_1$, and 
\noTrRL{s`yd} is the name of his father as 
the word \noTrRL{bn} (son of) indicates. 
The sequence of names from $n_1$ to $n_5$ 
constitutes the \noArRL{sanad} 
of the \noArRL{.hady_t}. 
The second part of the \noArRL{.hady_t} is the 
\RL{matn} (content) and constitutes the actual content 
and tradition.

@@

In this paper, we present the hadith {narrator chain 
extractor using morphological analysis} (NaCEMA), 
a novel technique that automates
extracting the \noArRL{sanad} from the books of traditions 
into a diagram of chains of narrators. NaCEMA segments
a book into narrations, each narration into its \noArRL{matn} 
and its \noArRL{sanad}, 
and each sanad into its separate narrators. 
This work will lead to 
automate an exhaustive automated \noArRL{.hady_t} authentication 
effort in the future as we plan to 
automatically analyze the biography books.

@@

%Several researchers~\cite{Hadithopaedia:08} attempted to automate 
%the analysis of the hadith literature. 
In this paper, we use Sarf to successfully automate the
analysis of three books of hadith selected 
arbitrarily~\cite{IbnHanbal,AlKulayni,AlTousi}\footnote{We obtained
  the digitized books from online sources such as 
  \href{http://www.yasoob.com/}{http://www.yasoob.com/} and 
  \href{http://www.al-eman.com/}{http://www.al-eman.com/}. }
%  Our understanding is that those books are historical documents with 
 % public IP.}.
%\setcode{utf8}
%\begin{arabtext}
%حدثنا \emphasize{هشيم}، أخبرنا \emphasize{إسماعيل بن أبي خالد}،
% عن أبي إسحاق، عن سعيد بن جبير، قال كنا
%    مع ابن عمر رضي الله عنه حيث أفاض من
%    عرفات إلى جمع فصلى بنا المغرب ومضى
%\end{arabtext}
%\setcode{standard}
NaCEMA accepts a book as input
and segments it into a vector of hadiths. 
NaCEMA segments each hadith into its
sanad and matn parts. 
NaCEMA 
detects the chain of narrators in the sanad and 
the relation that links each narrator to his ancestor and 
predecessor in the chain. 
It also detects the full name of each narrator that is
composed of several proper names with connectors
in between.

@@

The directed acyclic graph (DAG) 
in Figure~\ref{f:narrators} shows a partial order relation (POR) 
between narrators.
We automatically extracted the POR from three books of hadith 
selected arbitrarily~\cite{IbnHanbal,AlKulayni,AlTousi}.
The nodes in boxes are the \RL{matn} of the hadith, 
and the other nodes are the narrators.
Given a set of chains of narrators similar to 
Figure~\ref{f:exhadith} we formed the DAG using graph algorithms 
that merged equal names in the chains. 
We compared the narrators together using a morphological
distance function. 
We considered each narrator to have several morphological 
components and we rewarded every morphological match with 
a small $\delta$. 
Then we apply a threshold $\tau$ to the sum of rewards
decide the equality of the two narrators. 
We omitted the details of the distance function for brevity.

@@

We ran our experiments using a dual core 2.66 Ghz 64-bit processor 
with 4GB of memory using a Linux operating system. 
We report our results in Table~\ref{t:hadithresallresults} on the 
three hadith books.
NaCEMA detected narrator chains and segmented the books 
correctly with an accuracy above 90\%
in all cases and across all the traditions in the three books.

The row for ``word count'' is an indicator of the size of the data.
The %numbers in 
rows ``names'', ``narrators'', and ``chains'' denote the total
number of the corresponding entities per book.
The row ``ignored names'' reports the number of names detected 
outside the chains and thus are irrelevant to
the application. 
%In addition, the rows ``names/narrator'' and ``narrators/chain'' 
%report the average
%number of names per narrator and 
%the average number of narrators per chain respectively.

@@

This partial order graph is instrumental to automate
localization of narrators in biography documents and
segmenting biography documents in the future.
Given that a biography discusses a narrator 
and mentions his professors and students,
we plan to segment the biographies with a graph 
coloring algorithm
that traverses the text and colors the POR whenever
a name is found.

>>>>

\begin{document}
\maketitle
\begin{abstract}
A hadith in Islamic literature is a narration from the 
prophet Mohammad related by multiple narrators.
Establishing the authenticity of a hadith is an important task
in Islamic studies. 
Based on that Islamic scholars issue rules that affect the life
of people around the world. 
The literature lacks an automated exhaustive mechanism of 
authentication checks. 
This is currently manual work prone to error and almost
impossible to complete in the life time of a scholar due to the
large literature available. 
In this paper, we present a tool that uses 
an Arabic application-specific morphological analyzer 
to successfully automate the
analysis of three books of hadith selected 
arbitrarily and abstract 
each book into a graph of narrators related to each other. 
\end{quote}
\end{abstract}

\section{Introduction}
\novocalize

\transfalse
\begin{figure}[tb]
\center{
\resizebox{.9\columnwidth}{!}
{ \input{figs/exhadith.pdftex_t}}
\caption{Hadith abstraction example.}
\label{f:exhadith}
}
\end{figure}
% the FSMs for the three words
\transtrue
\vocalize

A \RL{.hady_t}~\footnote{In this document, 
we use the default ArabTeX transliteration style ZDMG.}
is a narration related to the prophet Mohammad
through a \RL{sanad} or a sequence of narrators. 
The collections of traditions are the second source of
jurisprudence after the \RL{qor'An} for all Islamic schools of thought. 
Figure~\ref{f:exhadith} shows an example \noArRL{.hady_t} in 
Arabic with its transliteration and translation. 
We show proper names in boxes connected
to form complex names of narrators. 
For example, 
\novocalize
\noTrRL{qtybT} is the first name 
of narrator $n_1$, and 
\noTrRL{s`yd} is the name of his father as 
the word \noTrRL{bn} (son of) indicates. 
The sequence of names from $n_1$ to $n_5$ 
constitutes the \noArRL{sanad} 
of the \noArRL{.hady_t}. 
The second part of the \noArRL{.hady_t} is the 
\RL{matn} (content) and constitutes the actual content 
of the tradition.

@@

In this paper, we present the hadith {narrator chain 
extractor using morphological analysis} (NaCEMA), 
a novel technique that automates
extracting the \noArRL{sanad} from the books of traditions 
into a diagram of chains of narrators. NaCEMA segments
a book into narrations, each narration into its \noArRL{matn} 
and its \noArRL{sanad}, 
and each sanad into its separate narrators. 
This work will lead to 
automate an exhaustive \noArRL{.hady_t} authentication 
effort in the future as we plan to 
automatically analyze the biography books.

@@

%Several researchers~\cite{Hadithopaedia:08} attempted to automate 
%the analysis of the hadith literature. 
In this paper, we use Sarf to successfully automate the
analysis of three books of hadith selected 
arbitrarily~\cite{IbnHanbal,AlKulayni,AlTousi}\footnote{We obtained
  the digitized books from online sources such as 
  \href{http://www.yasoob.com/}{http://www.yasoob.com/} and 
  \href{http://www.al-eman.com/}{http://www.al-eman.com/}. }.
%  Our understanding is that those books are historical documents with 
 % public IP.}.
%\setcode{utf8}
%\begin{arabtext}
%حدثنا \emphasize{هشيم}، أخبرنا \emphasize{إسماعيل بن أبي خالد}،
% عن أبي إسحاق، عن سعيد بن جبير، قال كنا
%    مع ابن عمر رضي الله عنه حيث أفاض من
%    عرفات إلى جمع فصلى بنا المغرب ومضى
%\end{arabtext}
%\setcode{standard}
NaCEMA accepts a book as input
and segments it into a vector of hadiths. 
NaCEMA segments each hadith into its
sanad and matn parts. 
NaCEMA 
detects the chain of narrators in the sanad and 
the relation that links each narrator to his ancestor and 
predecessor in the chain. 
It also detects the full name of each narrator that is
composed of several proper names with connectors
in between.

@@

The directed acyclic graph (DAG) 
in Figure~\ref{f:narrators} shows a partial order relation (POR) 
between narrators.
We automatically extracted the POR from the three books of 
hadith~\cite{IbnHanbal,AlKulayni,AlTousi}.
The nodes in boxes are the \RL{matn} of the hadith, 
and the other nodes are the narrators.
Given a set of chains of narrators similar to 
Figure~\ref{f:exhadith} we formed the DAG using graph algorithms 
that merged equal names in the chains. 
We compared the narrators together using a morphological
distance function. 
We considered each narrator to have several morphological 
components and we rewarded every morphological match with 
a small $\delta$. 
Then we applied a threshold $\tau$ to the sum of rewards
and decided the equality of the narrators. 
We omitted the details of the distance function for brevity.

@@

We ran our experiments using a dual core 2.66 Ghz 64-bit processor 
with 4GB of memory using a Linux operating system. 
We report our results in Table~\ref{t:hadithresallresults} on the 
three hadith books.
NaCEMA detected narrator chains and segmented the books 
correctly with a chain accuracy above 97\% and a narrator name
accuracy above 90\%
in all cases and across all the 1986 traditions in the three 
different books with different formats.

The row ``word count'' is an indicator of the size of the data.
The %numbers in 
rows ``names'', ``narrators'', and ``chains'' denote the total
number of the corresponding entities per book.
The row ``ignored names'' reports the number of names detected 
outside the chains and thus are irrelevant to
the application. 
%In addition, the rows ``names/narrator'' and ``narrators/chain'' 
%report the average
%number of names per narrator and 
%the average number of narrators per chain respectively.

@@

This partial order graph is instrumental to automate
localization of narrators in biography documents and
segmenting biography documents in the future.
Given that a biography discusses a narrator 
and mentions his professors and students,
we plan to segment the biographies with a graph 
coloring algorithm
that traverses the text and colors the POR node whenever
a name is found.
